---
title: "Lab 02 - Plastic waste"
author: "Anahatt Virk"
date: "06/20/25"
output: github_document
---

## Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
plastic_waste <- read.csv("data/plastic-waste.csv")
```

## Exercises

### Exercise 1

```{r plastic-waste-continent}
ggplot(data = plastic_waste, aes(x = plastic_waste_per_cap)) +
  geom_histogram(binwidth = 0.2) +
  facet_wrap(~continent) +
  labs(
title = "Plastic Waste per Capita by Continent", x = "Plastic Waste per Capita", y = "Continent"
    )
```

Based on the histograms plotted for plastic waste per capita by continent, it can be seen that Europe and North America produce the highest amount. This indicates that plastic waste per capita is likely higher in more industrialized continents. 

### Exercise 2

```{r plastic-waste-density}
ggplot(data = plastic_waste, mapping = aes(x = plastic_waste_per_cap, color = continent,
    fill = continent)) +
  geom_density(alpha = 0.2) +
  labs (title = "Density Plot of Plastic Waste per Capita by Continent", x = "Plastic Waste per Capita", y = "Density")
```

Color and fill are defined by the mapping aesthetics of the plot because they are dependent on the variable "continent." On the other hand, the alpha level is the same for all continents and is not linked to a variable.  

### Exercise 3

```{r plastic-waste-violin}
ggplot(data = plastic_waste, mapping = aes(x = continent, y = plastic_waste_per_cap)) +
  geom_violin()
```

Violin plots are useful in order to show the full distribution of the data. It can be easily seen if the data clusters or if there are outliers present. While violin plots provide more information about the shape of the data, box plots are helpful when a clearer view of the summary statistics is needed. Here, the median, quartiles, and the interquartile range are clearly displayed. Outliers are also  highlighted as individual points, which violin plots do not do. 

### Exercise 4

Remove this text, and add your answer for Exercise 4 here.

```{r plastic-waste-mismanaged}
# insert code here
```

```{r plastic-waste-mismanaged-continent}
# insert code here
```

```{r plastic-waste-population-total}
# insert code here
```

```{r plastic-waste-population-coastal}
# insert code here
```

### Exercise 5

Remove this text, and add your answer for Exercise 5 here.

```{r recreate-viz}
# insert code here
```


